%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Relevant & Associated Publications:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \singlespacing  %set single line spacing
\chapter*{Acknowledgements}
%\section*{Acknowledgement}

%\section*{Acknowledgement}

To all the teachers who have shaped me, this achievement would not have been possible without your guidance and constant support. Among them, I would like to express my deepest gratitude to Dr. Paul Orkwis (Dr. O), my mentor, who has stood by me through thick and thin. This project was his long-held dream, and I am honored to have played a role in bringing it to fruition.

I extend my sincere appreciation to my dissertation committee: Dr. Daniel Cuppoletti (Chair), Dr. Paul Orkwis, Dr. Prashant Khare, and Dr. Ephraim Gutmark. Your invaluable feedback, guidance, and support have shaped this dissertation, and I am truly grateful for the time and effort you dedicated to my work.

To Dr. Daniel Cuppoletti (Dan), thank you for your energy and support, especially towards the end. Your guidance helped me shape a new perspective on this work. Our discussions spawned ideas of particle dynamics history (PDH), case studies, and essential analytical tools, along with many brilliant insights. Your support was instrumental in helping me achieve the final vision of this work.

Dr. Prashant Khare has been a constant presence behind the scenes, providing invaluable support to me and many UC Aerospace students. Your ideas have shaped UC’s computing resources and influenced a significant part of this work. Thank you for your service.

To Dr. Ephraim Gutmark, thank you for your wisdom in shaping my research. Your insights helped refine the tone of my dissertation, and I sincerely appreciate your collaboration despite your many commitments.

I would also like to thank Dr. Mark Turner for his time at UC and for always making time for me when needed. I thank my collaborators from Florida State University (FSU), Dr. Rajan Kumar, Dr. Jonas Gustavsson, and Kevin Hernandez-Lichtl, for recognizing my work and allowing me to test the tools I have developed on their supersonic wind tunnel. Thank you to Dr. Sivaram Gogeneni, who took an interest in my work and progress with quarterly updates and provided me with essential contacts that helped me hone this work.

Many thanks to the Aerospace Department and Graduate School staff for their support. I am also grateful to my colleagues, Lana Petrou and Tim McGeorge, from my graduate assistantship at OCC, who continue to celebrate my personal and professional achievements. I would also like to thank the fantastic staff at Advanced Research Computing (ARC) for their constant support in enabling me to gain the knowledge I needed to focus on my research. Thanks to the Armstrong Institute for Space, Technology, and Research (ASTRO) grant for providing the funds required to support this work.

I am grateful to many of my old and new friends who have supported me during my time at UC. I want to thank Dr. Aatresh Karnam and Kranthi Yellugari from the Gas Dynamics and Propulsion Laboratory for their long-time collaboration and support during my many phases of graduate education. I also would like to thank many of my lab mates, especially Justin Holder, who has been a continuous emotional and professional support during the long hours we spent in the lab together, and Dr. Matteo Ugolotti, who provided me with a lot of collaboration and helped me navigate this academic world. I will surely miss our time together.

To my Amma, Nanna, and Baby, along with Ranga and Ayra, words are not enough to express my gratitude. Your love, sacrifices, and encouragement have been the foundation of my journey. This achievement is as much yours as it is mine.

Finally, I would like to express my love and gratitude to my love, Rasika Patil, who has always been my source of confidence and strength. Your unwavering support has made this work possible. I look forward to a long and fulfilling journey together, where we continue to uplift each other both personally and professionally while dedicating ourselves to making a meaningful impact on society.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
