\cleardoublepage
\phantomsection %to correct hyperlink reference issue
\addcontentsline{toc}{chapter}{Publications}
\chapter*{Publications}\label{ch:publications}
Below are the publications from the current work:
\begin{enumerate}
    \item D. Kalagotla, P. D. Orkwis, and M. Turner, \textit{Methodology to study the behavior of tracer particles in the flow field of rotor 37 using CFD data}, Conference Paper, 2020
    \item D. Kalagotla, A. Karnam, and E. J. Gutmark, \textit{Comparison of flow characteristics of single and twin rectangular jets using overflow code}, Conference Paper, 2020.
    \item H. Chhabra, D. Kalagotla, P. D. Orkwis, \textit{Python based API to post-process CFD data}, Conference Paper, 2023.
    \item D. Kalagotla, P. D. Orkwis, \textit{An Object-Oriented Approach to Tracking Particles in a Flow}, Conference Paper, 2023.
    \item D. Kalagotla, P. D. Orkwis, \textit{A New Approach to Synthetic PIV Data Based on Particle Dynamics History}, Conference Paper, 2024.
    \item D. Kalagotla, P. D. Orkwis, \textit{Numerical Analysis of Drag Models and Particle Size Estimation for PIV in Supersonic Flows}, Conference Paper, 2024.
    \item D. Kalagotla, P. D. Orkwis, D. Cuppoletti, \textit{Deep Learning based Particle Inertia Bias Corrector for Supersonic PIV}, Conference Paper, 2025.
    \item D. Kalagotla, K. Hernandez-Lichtl, J. Gustavsson, R. Kumar, D. Cuppoletti, P. Orkwis, \textit{Particle Size Distribution Estimation in Supersonic PIV Experiments}, Journal Paper (Submitted to AIAAJ), 2025.
    \item D. Kalagotla, D. Cuppoletti, P. Orkwis, \textit{Deep Learning Based Particle Inertia Bias Reduction for Shock-Dominated PIV Data}, Journal Paper (In Progress), 2025.
    \item D. Kalagotla, D. Cuppoletti, P. D. Orkwis, \textit{On the Effect of Stokes Number in Studying Complex Supersonic Flows Using PIV}, Journal Paper (In Progress), 2025.
    

\end{enumerate}